---
title: "Run Pixy and visualize results"
format:
  html:
    code-fold: show
    code-tools: true
toc: true
toc-title: Document Contents
number-sections: true
embed-resources: true
---

### Step 1: I calculated genetic diversity in 1Mb sliding using the depth-filtered, all-sites vcf as input for each sample. I did this using Pixy, by running the following code on the UCLA cluster:
```{bash, eval=FALSE}
#### submit_job.sh START ####
#!/bin/bash
#$ -cwd
#$ -o ./joblog.$JOB_ID.$TASK_ID.txt          #set the job log output file
#$ -j y                   #set error = Merged with joblog
#$ -l h_rt=10:00:00,h_data=5G   #specify requested resources (h_rt gives time request in 'hrs:mins:secs' format) (h_data specifies requested RAM per task) (highp=TRUE means run it on Aguillon Lab owned nodes)
#$ -pe shared 1              #specify number of CPUs requested
#  Job array indexes (1-25: by 1)
#$ -t 1-23:1

#load modules
. /u/local/Modules/default/init/modules.sh
#load programs
module load conda
conda activate pixy2 #version (2.3.2)
module load bcftools #(version: 1.11)

#base name so that the array gets one sample per job
basename_array=$( head -n${SGE_TASK_ID} basenames.txt | tail -n1 )

#write pop file for this sample
echo -e "${basename_array}\tpop1" > ${basename_array}_popfile.txt

#run pixy in 1Mb sliding windows
### use "--chromosomes 'X'" to restrict to just certain chromosomes in the vcf. Default is all chromosomes in vcf
pixy --stats pi \
--vcf /u/home/d/dderaad/project-aguillon/eagle.mountain.scrubjays/bwa.mem/vcf/${basename_array}.depth.filtered.vcf.gz \
--populations ${basename_array}_popfile.txt \
--window_size 1000000 \
--n_cores 1 \
--output_folder /u/home/d/dderaad/project-aguillon/eagle.mountain.scrubjays/main.text/pixy \
--output_prefix ${basename_array}

```


### Then, read in the output file for each sample, and create a single, long-format table to hold all of this information
```{r}
#Get a list of all the output files (.posterior file extension)
pixy_files<-Sys.glob("~/Desktop/2026.eagle.mountain/pixy/*_pi.txt")

#open a dataframe to hold all data together
complete.df<-c()

#in a loop, read in each dataframe, rename according to that sample, and combine
for (i in 1:length(pixy_files)){
  #read in dataframe
  df<-read.table(pixy_files[i], header = T)
  df$pop<-gsub("_pi.txt","",gsub("/Users/devonderaad/Desktop/2026.eagle.mountain/pixy/","",pixy_files[i]))
  complete.df<-rbind.data.frame(complete.df,df)
}

#check this worked
table(complete.df$pop)
```

### Plot genetic diversity along chrom1
```{r}
#load libraries
library(dplyr)
library(ggplot2)
library(gghalves)

#plot in loop
for (i in 1:length(unique(complete.df$pop))){
  print(
# plotting summary statistics along a single chromosome
complete.df %>%
  filter(chromosome == "NC_091013.1") %>%
  filter(pop == unique(complete.df$pop)[i]) %>%
  mutate(chr_position = ((window_pos_1 + window_pos_2)/2)/1000000) %>%
  ggplot(aes(x = chr_position, y = avg_pi*1000))+
  geom_col(fill="black")+
  xlab(paste("Position on Chromosome 1 (Mb) -",unique(complete.df$pop)[i]))+
  ylab("Hets per Kb")+
  theme_bw()+
  scale_x_continuous(expand = c(0, 0))+
  scale_y_continuous(expand = c(0, 0))
)
}
```

### Plot genetic diversity along chrom2
```{r}
for (i in 1:length(unique(complete.df$pop))){
  print(
# plotting summary statistics along a single chromosome
complete.df %>%
  filter(chromosome == "NC_091015.1") %>%
  filter(pop == unique(complete.df$pop)[i]) %>%
  mutate(chr_position = ((window_pos_1 + window_pos_2)/2)/1000000) %>%
  ggplot(aes(x = chr_position, y = avg_pi*1000))+
  geom_col(fill="black")+
  xlab(paste("Position on Chromosome 2 (Mb) -",unique(complete.df$pop)[i]))+
  ylab("Hets per Kb")+
  theme_bw()+
  scale_x_continuous(expand = c(0, 0))+
  scale_y_continuous(expand = c(0, 0))
)
}
```

### Plot genetic diversity along chrom4
```{r}
for (i in 1:length(unique(complete.df$pop))){
  print(
# plotting summary statistics along a single chromosome
complete.df %>%
  filter(chromosome == "NC_091017.1") %>%
  filter(pop == unique(complete.df$pop)[i]) %>%
  mutate(chr_position = ((window_pos_1 + window_pos_2)/2)/1000000) %>%
  ggplot(aes(x = chr_position, y = avg_pi*1000))+
  geom_col(fill="black")+
  xlab(paste("Position on Chromosome 4 (Mb) -",unique(complete.df$pop)[i]))+
  ylab("Hets per Kb")+
  theme_bw()+
  scale_x_continuous(expand = c(0, 0))+
  scale_y_continuous(expand = c(0, 0))
)
}
```

### Calc genome-wide genetic diversity per sample
```{r}
#open a new dataframe to store this information
gendiv.df<-data.frame(samp=character(),
                      gendiv=numeric())

#using a for loop, calculate genome-wide genetic diversity in each sample and store it in a dataframe
for (i in 1:length(unique(complete.df$pop))){
  
  ### isolate the given sample
  x <- complete.df[complete.df$pop == unique(complete.df$pop)[i],]
  
  #exclude the Z, W, and mtDNA scaffolds:
  x<-x[x$chromosome != "NC_138418.1" & x$chromosome != "NC_138417.1" & x$chromosome != "NC_051467.1",]
  
  #calculate the proportion of het sites across all called sites for this genome
  #store in df alongside sample info
  gendiv.df[i,]<-c(unique(complete.df$pop)[i],sum(x$count_diffs, na.rm = TRUE) / sum(x$no_sites, na.rm = TRUE))
}
```


### Plot genome-wide genetic diversity per sample by subspecies
```{r}
#make numeric
gendiv.df$gendiv<-as.numeric(gendiv.df$gendiv)

#add subspecies info to data frame
gendiv.df$subspecies<-c(rep("cali", times=3),rep("obsc", times=6),"hypo",rep("obsc", times=3),rep("cana", times=3),"gris","gris",rep("nev", times=3),"gris","wood")

#add sample type info
gendiv.df$type<-c(rep("muscle", times=10),rep("toepad", times=8),rep("muscle", times=3),rep("toepad", times=2))

#reorder factors for plotting
gendiv.df$subspecies <- factor(gendiv.df$subspecies, levels = unique(gendiv.df$subspecies))

#plot
ggplot(gendiv.df, aes(x = subspecies, y = gendiv*1000, shape=type)) +
  geom_half_boxplot(side = "l", outlier.shape = NA, errorbar.draw = FALSE) +
  geom_half_point(side = "r", transformation = position_jitter(width = 0.05, height = 0)) +
  labs(y = "Heterozygous sites per Kb") +
  theme_classic()

#replot
ggplot(gendiv.df, aes(x = subspecies, y = gendiv*1000)) +
  geom_half_boxplot(side = "l", outlier.shape = NA, errorbar.draw = FALSE) +
  geom_half_point(side = "r", transformation = position_jitter(width = 0.05, height = 0)) +
  labs(y = "Heterozygous sites per Kb") +
  theme_classic()

#save
plotx<-ggplot(gendiv.df, aes(x = subspecies, y = gendiv*1000)) +
  geom_half_boxplot(side = "l", outlier.shape = NA, errorbar.draw = FALSE) +
  geom_half_point(side = "r", transformation = position_jitter(width = 0.05, height = 0)) +
  labs(y = "Heterozygous sites per Kb") +
  theme_classic()

#ggsave("~/Desktop/2026.eagle.mountain/gendiv.pdf", plotx, width = 6.5, height = 2, units = "in")
```

